---
title: "Group Project Cleaning File"
author: "Connor Harrison, Dong Hoon Lee, Haorui Sun"
date: "April 8, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Purpose
For our project, we are using the "" data set from the .....Before we can perform any anlysis, we need to import and clean the data into a usable format. This will include retaining only the variables we place to use, transforming variable coding, and controlling for missing values. 


## Import Data Set
```{r}
library(haven)
epcg17 <- read_dta("epcg17.dta")
View(epcg17)
```

## Relevant Variables
This data set is from a large, comprehensive survey. As such, it contains data on a wide variety of topicsm many of which are not relevant for our analysis. This section of code retains only the variables that we will explore 
```{r}

## selecting variables

epcg17_new <-  epcg17 %>% 
  filter(., age<36) %>% 
  # selecting age, race, undergrad loan variable
  select(age, racem, ugloanr, 
         #primary/secondary work activity, public/private status of undergrad institution, year ba received
         actcap:acttch, bapbpr, bayr,
         # reason for changing employer
         chchg:chloc, chot:chsch,
         # visa type for non-US citizen, highest degree type, year highest degree awarded, father and mother's education level
         ctzfor, dgrdg, dgryr, eddad, edmom,
         #region code for employer, detailed employer sector, 
         emrg, emsecdt, facadv:facsoc)
```

