---
title: "Group Project Cleaning File"
author: "Connor Harrison, Dong Hoon Lee, Haorui Sun"
date: "April 8, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Purpose
For our project, we are using the "" data set from the .....Before we can perform any anlysis, we need to import and clean the data into a usable format. This will include retaining only the variables we place to use, transforming variable coding, and controlling for missing values. 


## Import Data Set
```{r}
library(haven)
library(tidyverse)
epcg17 <- read_dta("epcg17.dta")
View(epcg17)
```

## Relevant Variables
This data set is from a large, comprehensive survey. As such, it contains data on a wide variety of topicsm many of which are not relevant for our analysis. This section of code retains only the variables that we will explore 
```{r}

## filtering by age, creating degree variables, and selecting other variables

epcg17_new <-  epcg17 %>% 
  filter(., age<36) %>% 
  # selecting age, race, gender, undergraduate loan amoutn, graduate loan amount, physical disability indicator
  select(age, racethm, ugloanr, gender, grloanr, hcapin,
         #degree information, second to fifth highest degrees, highest degree, most recent degree
         d2dg, d3dg, d4dg, d5dg, dgrdg, mrdg,
         #primary/secondary work activity, public/private status of undergrad institution, year ba received
         actcap:acttch, bapbpr, bayr,
         # reason for changing employer
         chchg:chloc, chot:chsch,
         # visa type for non-US citizen, year highest degree awarded, father and mother's education level
         ctzfor, dgryr, eddad, edmom,
         #region code for employer, detailed employer sector, 
         emrg, emsecdt, facadv:facsoc,
         # hours per week worked, job benefits, job satisfaction, labor force status, looking for work, marital status
         hrswk, jobins:jobproft, jobvac, jobsatis, lfstat, lookwk, marsta,
         # job code for last job, job code for principal job, field of study of BA degree, employer type
         n2ocmlst, n2ocprmg, nbamed, nedtp,
         # reasons for working outside field of highest degree, most important reason for working outside field
         nrchg:nrpay, nrrea,
         # reasons for not working, extent job is related to highest degree, reason worked part-time
         nwfam:nwstu, ocedrlp, pjfam:pjstu,
         # respondent location, salary, satisfaction in prinicipal job for..., spouse working indicator, year principal job started
         resploc, salary, satadv:satsoc, spowk, strtyr,
         # amount still owed from undergrad loan, amount still owed from grad loan, veteran status 
         ugower, grower, vetstat) %>% 
  # creating gender and race indicators
  mutate(., female = ifelse(gender=="F", 1,0)) %>% 
  mutate(., race_white = ifelse(racethm=="5",1,0 )) %>% 
  mutate(., race_black = ifelse(racethm=="3",1,0 )) %>% 
  mutate(., race_hispanic = ifelse(racethm=="4",1,0 )) %>%
  mutate(., race_asian = ifelse(racethm=="1",1,0 )) %>%
  mutate(., race_other = ifelse(racethm=="2" | racethm=="6" | racethm=="7", 1,0)) %>% 
  # replacing 9999998 to NA for salary variable
  mutate(., salary= ifelse(salary==9999998, NA, salary)) %>% 
  # bachelor degree indicator
  mutate(., bachelor_degree= ifelse(mrdg=="1", 1,
                                    ifelse(dgrdg=="1", 1,
                                           ifelse(d2dg=="1", 1,
                                                  ifelse(d3dg=="1", 1,
                                                         ifelse(d4dg=="1",1,
                                                                ifelse(d5dg=="1", 1, 0))))))) %>%
  # master's degree indicator
  mutate(., master_degree= ifelse(mrdg=="2", 1,
                                    ifelse(dgrdg=="2", 1,
                                           ifelse(d2dg=="2", 1,
                                                  ifelse(d3dg=="2", 1,
                                                         ifelse(d4dg=="2",1,
                                                                ifelse(d5dg=="2", 1, 0))))))) %>%
  mutate(., doctorate_degree= ifelse(mrdg=="3", 1,
                                    ifelse(dgrdg=="3", 1,
                                           ifelse(d2dg=="3", 1,
                                                  ifelse(d3dg=="3", 1,
                                                         ifelse(d4dg=="3",1,
                                                                ifelse(d5dg=="3", 1, 0))))))) %>%
  # professional degree
  mutate(., professional_degree= ifelse(mrdg=="4", 1,
                                    ifelse(dgrdg=="4", 1,
                                           ifelse(d2dg=="4", 1,
                                                  ifelse(d3dg=="4", 1,
                                                         ifelse(d4dg=="4",1,
                                                                ifelse(d5dg=="4", 1, 0)))))))

head(epcg17_new)
```

