---
title: "Group Project Cleaning File"
author: "Connor Harrison"
date: "April 8, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Purpose
For our project, we are using the "" data set from the .....Before we can perform any anlysis, we need to import and clean the data into a usable format. This will include retaining only the variables we place to use, transforming variable coding, and controlling for missing values. 


## Import Data Set
```{r}
library(haven)
epcg17 <- read_dta("epcg17.dta")
View(epcg17)
```

## Relevant Variables
This data set is from a large, comprehensive survey. As such, it contains data on a wide variety of topicsm many of which are not relevant for our analysis. This section of code retains only the variables that we will explore 
```{r}
# Drop all variables that begin with "D". These are for degrees earned after the 1st.

# Keep all "E" variables; these are employment indicators

# Keep all "FA" variables; these are 'importance of jobs' questions

# Keep to undergraduate degrees only
```

