---
title: "Predicting Job Mismatch from Job Satisfaction"
author: "Harrison-Lee-Sun"
date: "2019/4/29"
output:
  slidy_presentation:
    incremental: yes
  ioslides_presentation:
    incremental: yes
    widescreen: yes
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Problem statement and background

- The cost of higher education has increased but incomes hasn't

- Student debt has also increased significantly

- Student loan debt in 2018: $1.5 trillion

- Graduating class of 2016 has average of $37,000 student loan.

## Problem statement and background

- Studies have shown that student debt impacts life decisions

- Student debt may pressure and lead to job mismatches

- Hypothesis: high levels of student loan debt have negative impacts on employment outcomes

## Research design and approach

- Goals

1. Data visualization: to test any trends or relationships between variables of interest

2. Machine learning: to find the best predictor for the overall job satisfaction

## Methods / Tools

- Dataset: National Survey of College Graduates (NSCG)

- Data Visualization 
1. Job satisfaction over Loan Amounts (Region and Race)
2. Comparison of job satisfaction aspects
3. Distribution of job satisfaction responses by salary and age groups

## Methods / Tools

- Machine Learning 
1. Job Satisfaction - Recent Graduates
2. Job Satisfaction - Full Sample
3. Salary Satisfaction
4. Career Advancement Satisfaction

- 6 Models: K-Nearest Neighbor, Regression Tree, Random Forest
  -With and without tuning parameters

## Results - Visualization 1

```{r, include=FALSE}
# Load Packages
library(haven)
library(tidyverse)
library(caret)
library(recipes)
library(pdp)
library(skimr)
library(foreach)
library(pastecs)
library(ggplot2)

## Data Wrangling - Won't be included in the knit pdf output.
## filtering by age, selecting variables of interest, creating key indicators
epcg17 <- read_dta("epcg17.dta")
epcg17_new <-  epcg17 %>% 
  filter(., age<36 & bast=="099" & bayr>1975 & ugloanr!="01" & emrg<10) %>% 
  # selecting age, race, gender, undergraduate loan amoutn, graduate loan amount, physical disability indicator
  select(age, racethm, ugloanr, gender, grloanr, hcapin,
         #degree information, second to fifth highest degrees, highest degree, most recent degree
         d2dg, d3dg, d4dg, d5dg, dgrdg, mrdg,
         #primary/secondary work activity, public/private status of undergrad institution, year ba received
         actcap:acttch, bapbpr, bayr,
         # reason for changing employer
         chchg:chloc, chot:chsch,
         # visa type for non-US citizen, year highest degree awarded, father and mother's education level
         ctzfor, dgryr, eddad, edmom,
         #region code for employer, employer type, employer size, detailed employer sector, 
         emrg, emtp, emsize, emsecdt, emsecsm, facadv:facsoc,
         # hours per week worked, job benefits, job satisfaction, labor force status, looking for work, marital status
         hrswk, jobins:jobproft, jobvac, jobsatis, lfstat, lookwk, marsta,
         # job code for last job, job code for principal job, field of study of BA degree, employer type
         n2ocmlst, n2ocprmg, nbamed, nedtp,
         # reasons for working outside field of highest degree, most important reason for working outside field
         nrchg:nrpay, nrrea,
         # reasons for not working, extent job is related to highest degree, reason worked part-time
         nwfam:nwstu, ocedrlp, pjfam:pjstu,
         # respondent location, salary, satisfaction in prinicipal job for..., spouse working indicator, year principal job started
         resploc, salary, satadv:satsoc, spowk, strtyr,
         # amount still owed from undergrad loan, amount still owed from grad loan, veteran status 
         ugower, grower, vetstat) %>% 
  # creating gender and race indicators
  mutate(., female = ifelse(gender=="F", 1,0)) %>% 
  mutate(., race_white = ifelse(racethm=="5",1,0 )) %>% 
  mutate(., race_black = ifelse(racethm=="3",1,0 )) %>% 
  mutate(., race_hispanic = ifelse(racethm=="4",1,0 )) %>%
  mutate(., race_asian = ifelse(racethm=="1",1,0 )) %>%
  mutate(., race_other = ifelse(racethm=="2" | racethm=="6" | racethm=="7", 1,0)) %>% 
  # replacing 9999998 to NA for salary variable
  mutate(., salary= ifelse(salary==9999998, NA, salary)) %>% 
  # bachelor degree indicator
  mutate(., bachelor_degree= ifelse(mrdg=="1", 1,
                                    ifelse(dgrdg=="1", 1,
                                           ifelse(d2dg=="1", 1,
                                                  ifelse(d3dg=="1", 1,
                                                         ifelse(d4dg=="1",1,
                                                                ifelse(d5dg=="1", 1, 0))))))) %>%
  # master's degree indicator
  mutate(., master_degree= ifelse(mrdg=="2", 1,
                                    ifelse(dgrdg=="2", 1,
                                           ifelse(d2dg=="2", 1,
                                                  ifelse(d3dg=="2", 1,
                                                         ifelse(d4dg=="2",1,
                                                                ifelse(d5dg=="2", 1, 0))))))) %>%
  mutate(., doctorate_degree= ifelse(mrdg=="3", 1,
                                    ifelse(dgrdg=="3", 1,
                                           ifelse(d2dg=="3", 1,
                                                  ifelse(d3dg=="3", 1,
                                                         ifelse(d4dg=="3",1,
                                                                ifelse(d5dg=="3", 1, 0))))))) %>%
  # professional degree
  mutate(., professional_degree= ifelse(mrdg=="4", 1,
                                    ifelse(dgrdg=="4", 1,
                                           ifelse(d2dg=="4", 1,
                                                  ifelse(d3dg=="4", 1,
                                                         ifelse(d4dg=="4",1,
                                                                ifelse(d5dg=="4", 1, 0))))))) %>%
  mutate(.,satadv_satisfied = if_else(satadv=="1" | satadv =="2", 1, 0)) %>% 
  mutate(.,satben_satisfied = if_else(satben=="1" | satben =="2", 1, 0)) %>%
  mutate(.,satsal_satisfied = if_else(satsal=="1" | satsal =="2", 1, 0)) %>%
  mutate(.,satsec_satisfied = if_else(satsec=="1" | satsec =="2", 1, 0)) %>%
  mutate(.,satresp_satisfied = if_else(satresp=="1" | satresp =="2", 1, 0)) %>% 
  mutate(.,job_satisfied = if_else(jobsatis=="1" | jobsatis =="2", 1, 0))
```

```{r}
## Visual 2
# data set creation for the visual 2
dataforplot2 <- epcg17_new %>%
  filter(., jobsatis!="L") %>%
  select(., ugloanr,satadv_satisfied:satresp_satisfied,job_satisfied,emrg) %>% 
  mutate(.,obscount=1) %>%
  mutate(.,ugloanr_group=ifelse(as.numeric(ugloanr)==2, 1,
                                ifelse(as.numeric(ugloanr)>2 & as.numeric(ugloanr)<6,2,
                                       ifelse(as.numeric(ugloanr)>5 & as.numeric(ugloanr)<9, 3,
                                              ifelse(as.numeric(ugloanr)>8 &  as.numeric(ugloanr)<12,4,5))))) %>%
  group_by(.,ugloanr_group, emrg) %>% 
  summarize(.,mean_job_sat=mean(job_satisfied),
            obscount=sum(obscount)) %>% 
  mutate(.,emrg_char=ifelse(emrg=="01","New England",
                            ifelse(emrg=="02", "Middle Atlantic",
                                   ifelse(emrg=="03", "East North Central",
                                          ifelse(emrg=="04", "West North Central",
                                                 ifelse(emrg=="05", "South Central",
                                                        ifelse(emrg=="06", "East South Central",
                                                               ifelse(emrg=="07", "West South Central",
                                                                      ifelse(emrg=="08", "Mountain", "Pacific and US Territories"))))))))) %>% 
  filter(., obscount>=50) %>% 
  mutate(.,emrg_factor=factor(emrg_char, c("New England", "Middle Atlantic", "East North Central", "West North Central",
                                            "South Central", "East South Central", "West South Central",
                                            "Mountain", "Pacific and US Territories"))) %>%
  mutate(.,mean_job_sat_100 = mean_job_sat * 100)

# ggplot code for visual 2
ggplot(dataforplot2, aes(x=emrg_factor, y=mean_job_sat_100, color=as.character(ugloanr_group), group=as.character(ugloanr_group))) +
  geom_point(size=3.5, alpha=.8) +
  geom_text(aes(label=round(mean_job_sat_100,0)), size=4, vjust = -.8, nudge_y = 0.05) +
  scale_color_manual(name = "Undergrad Loan Amount",
                     values = c("1"="yellowgreen", "2"="yellow3", "3"="orange1", "4"="orangered1", "5" = "red4"),
                     labels = c("1"="$0", "2"="$1-30,000", "3"="$30,001-60,000", "4"="$60,001-90,000", "5" = "More than $90,000")) +
  labs(title ="Percent Satisfied with Job by Undergraduate Loan Amount and Region",
       caption = "Data: National Survey of College Graduates") +
  coord_flip() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(),
        panel.grid.major.y =element_line(colour = "grey", size = 0.5))

```


## Results - Visualization 2
```{r}
## Results - Visualization
# Visual 4 data preparation
dataforplot4 <- epcg17_new %>%
  filter(., jobsatis!="L") %>%
  select(., ugloanr,emrg, satadv_satisfied:satresp_satisfied, job_satisfied) %>% 
  mutate(.,obscount=1) %>%
  mutate(.,ugloanr_group=ifelse(as.numeric(ugloanr)==2, 1,
                                ifelse(as.numeric(ugloanr)>2 & as.numeric(ugloanr)<6,2,
                                       ifelse(as.numeric(ugloanr)>5 & as.numeric(ugloanr)<9, 3,
                                              ifelse(as.numeric(ugloanr)>8 & as.numeric(ugloanr)<12,4,5))))) %>%
  group_by(.,ugloanr_group, emrg) %>%
  summarize(., mean1=mean(satadv_satisfied),
            mean2=mean(satben_satisfied), mean3=mean(satsal_satisfied),
            mean4=mean(satsec_satisfied), mean5=mean(satresp_satisfied),
            mean6=mean(job_satisfied), obscount=sum(obscount)) %>%
  filter(., obscount>=50) %>% 
  ungroup(.,ugloanr_group, emrg) %>% 
  gather(., key, value,mean1:mean6) %>% 
  mutate(.,emrg_char=ifelse(emrg=="01","New England",
                            ifelse(emrg=="02", "Middle Atlantic",
                                   ifelse(emrg=="03", "East North Central",
                                          ifelse(emrg=="04", "West North Central",
                                                 ifelse(emrg=="05", "South Central",
                                                        ifelse(emrg=="06", "East South Central",
                                                               ifelse(emrg=="07", "West South Central",
                                                                      ifelse(emrg=="08", "Mountain", "Pacific and US Territories"))))))))) %>% 
  mutate(.,emrg_factor=factor(emrg_char, c("New England", "Middle Atlantic", "East North Central", "West North Central",
                                           "South Central", "East South Central", "West South Central",
                                           "Mountain", "Pacific and US Territories"))) %>% 
  mutate(.,satisfaction_type=ifelse(key=="mean1","Career Advancement",
                                    ifelse(key=="mean2", "Benefits",
                                           ifelse(key=="mean3", "Salary",
                                                  ifelse(key=="mean4", "Job Security",
                                                         ifelse(key=="mean5", "Responsibility", "Job - Overall")))))) %>% 
  mutate(.,mean100=value*100)

## Visual 4
ggplot(dataforplot4, aes(x=as.character(ugloanr_group),y=mean100, color=satisfaction_type, group=satisfaction_type)) +
  geom_point() +
  geom_line(size=1.3, alpha=.6) +
  facet_wrap(~emrg_factor) +
  scale_color_brewer(name="Satisfaction Category",
                     type="qual",
                     palette="Set1") +
  scale_x_discrete(labels = c("1"="$0", "2"="$1-30K", "3"="$30k-60k", "4"="$60k-90k", "5" = "More than $90k")) +
  scale_y_continuous(limits = c(60, 100),breaks = seq(60, 100, 10)) +
  labs(title ="Types of Job Satisfaction by Undergraduate Loan Amounts and by Region",
       caption = "Data: National Survey of College Graduates",
       y = "Percent Satisfied") +
  theme(axis.title.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(),
        panel.grid.major.y =element_line(colour = "grey", size = 0.5),
        axis.text.x = element_text(size=8, angle=20))
```

## Results - Visualization 3
```{r}
##Visual 9
dataforplot9_1 <- epcg17_new %>%
  filter(., jobsatis!="L") %>%
  mutate(.,salary_range=ifelse(salary==0, "$0",
                               ifelse(salary>0 & salary <=30000, "$1-30K",
                                      ifelse(salary>30000 & salary <=60000, "$30K-60K",
                                             ifelse(salary>60000 & salary <=100000, "$60K-100K",
                                                    ifelse(salary>100000 & salary <=150000, "$100K-150K",">$150K")))))) %>%
  mutate(.,job_verysatisfied=ifelse(jobsatis=="1",1,0)) %>% 
  mutate(.,job_satisfied=ifelse(jobsatis=="2",1,0)) %>% 
  mutate(.,job_dissatisfied=ifelse(jobsatis=="3",1,0)) %>% 
  mutate(.,job_verydissatisfied=ifelse(jobsatis=="4",1,0)) %>% 
  mutate(.,obscount=1) %>% 
  group_by(.,salary_range) %>% 
  summarize(., mean1=mean(job_verysatisfied),
            mean2=mean(job_satisfied),
            mean3=mean(job_dissatisfied),
            mean4=mean(job_verydissatisfied),
            obscount=sum(obscount)) %>% 
  gather(., key, value,mean1:mean4) %>% 
  mutate(.,satis_level=ifelse(key=="mean1","Very Satisfied",
                              ifelse(key=="mean2", "Satisfied",
                                     ifelse(key=="mean3", "Dissatisfied","Very Dissatisfied")))) %>%
  mutate(.,satis_level_factor=factor(satis_level,c("Very Satisfied", "Satisfied","Dissatisfied","Very Dissatisfied"))) %>% 
  mutate(.,salary_factor=factor(salary_range, c("$0", "$1-30K","$30K-60K","$60K-100K","$100K-150K",">$150K"))) %>% 
  mutate(.,value100=value*100) %>% 
  mutate(.,satis_type="Job - Overall Satisfaction")

dataforplot9_2 <- epcg17_new %>%
  filter(., jobsatis!="L") %>%
  mutate(.,salary_range=ifelse(salary==0, "$0",
                               ifelse(salary>0 & salary <=30000, "$1-30K",
                                      ifelse(salary>30000 & salary <=60000, "$30K-60K",
                                             ifelse(salary>60000 & salary <=100000, "$60K-100K",
                                                    ifelse(salary>100000 & salary <=150000, "$100K-150K",">$150K")))))) %>%
  mutate(.,sal_verysatisfied=ifelse(satsal=="1",1,0)) %>% 
  mutate(.,sal_satisfied=ifelse(satsal=="2",1,0)) %>% 
  mutate(.,sal_dissatisfied=ifelse(satsal=="3",1,0)) %>% 
  mutate(.,sal_verydissatisfied=ifelse(satsal=="4",1,0)) %>% 
  mutate(.,obscount=1) %>% 
  group_by(.,salary_range) %>% 
  summarize(., mean1=mean(sal_verysatisfied),
            mean2=mean(sal_satisfied),
            mean3=mean(sal_dissatisfied),
            mean4=mean(sal_verydissatisfied),
            obscount=sum(obscount)) %>% 
  gather(., key, value,mean1:mean4) %>% 
  mutate(.,satis_level=ifelse(key=="mean1","Very Satisfied",
                              ifelse(key=="mean2", "Satisfied",
                                     ifelse(key=="mean3", "Dissatisfied","Very Dissatisfied")))) %>%
  mutate(.,satis_level_factor=factor(satis_level,c("Very Satisfied", "Satisfied","Dissatisfied","Very Dissatisfied"))) %>% 
  mutate(.,salary_factor=factor(salary_range, c("$0", "$1-30K","$30K-60K","$60K-100K","$100K-150K",">$150K"))) %>% 
  mutate(.,value100=value*100) %>% 
  mutate(.,satis_type="Salary Satisfaction")

dataforplot9 <- bind_rows(dataforplot9_1, dataforplot9_2)
  

ggplot(dataforplot9,aes(x=salary_factor, y=value100,fill=satis_level_factor)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=round(value100,0)), position=position_stack(vjust=0.5)) +
  facet_wrap(~satis_type)+
  labs(title ="Figure 9. Job and Salary Satisfaction by Salary Groups",
       caption = "Data: National Survey of College Graduates",
       y = "Percent") +
  scale_fill_brewer(name="Satisfaction Level",
                    type="div",
                    palette="RdBu",
                    direction = -1) +
  theme(axis.title.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size=8, angle=20))

```

```{r, include=FALSE, cache=TRUE}
### Prep Machine Learning, not to be included in slide output ###

epcg_ml <- select(epcg17_new, "age"=age, "race"=racethm, "gender"=gender, "salary"=salary, 
                  "employer_size"=emsize, "employer_region"=emrg, "highest_degree"=dgrdg, 
                  "employer_sector"=emsecdt, "loan_amount"=ugloanr, "job_satisfaction"=job_satisfied)

# Ensure all factor levels are correct
epcg_ml$job_satisfaction <- factor(epcg_ml$job_satisfaction, 
                                      levels = c(1, 0), 
                                      labels = c("Satisfied", "Not_Satisfied"))
# Set seed for replicability
set.seed(123)  

# Partition data into training and test
index = createDataPartition(epcg_ml$job_satisfaction,p=.75,list=F) 
epcg_train = epcg_ml[index,] # Use 75% of the data as training data 
epcg_test = epcg_ml[-index,] # holdout 25% as test data 

# Continuous variables that may need redistributing: salary
convert_salary <- . %>% mutate(salary = log(salary+1))

epcg_train2 <- epcg_train %>%  convert_salary()
epcg_test2 <- epcg_test %>%  convert_salary()

# Prep the recipe
rcp <- 
  recipe(job_satisfaction~.,epcg_train2) %>% 
  step_dummy(all_nominal(),-all_outcomes()) %>% 
  step_range(salary, age) %>%
  step_knnimpute(all_predictors()) %>%
  prep()

# Bake the recipe
epcg_train3 <- bake(rcp,epcg_train2)
epcg_test3 <- bake(rcp,epcg_test2)

# Set seed for replicability
set.seed(123)  

# Partition the data into 5 equal folds
folds <- createFolds(epcg_train3$job_satisfaction, k = 5) 

# Apply the folds
sapply(folds,length)

# Setting the control conditions to compare results across different model specifications
control_conditions <- 
  trainControl(method='cv',
               summaryFunction = twoClassSummary,
               classProbs = TRUE, 
               index = folds)

# KKN Model
epgc_mod_knn <-
  train(job_satisfaction ~ .,
        data=epcg_train3,
        method = "knn",
        metric = "ROC",
        trControl = control_conditions)

# Tuned KNN Model
knn_tune = expand.grid(k = c(1,3,5,10))
epgc_mod_knn2 <-
  train(job_satisfaction ~ .,
        data=epcg_train3,
        method = "knn",
        metric = "ROC",
        tuneGrid = knn_tune,
        trControl = control_conditions)

# Regression Tree Model
epgc_mod_cart <-
  train(job_satisfaction ~ .,
        data=epcg_train3,  
        method = "rpart", 
        metric = "ROC", 
        trControl = control_conditions)

# Set tuning parameter for regression tree
tune_cart2 <- expand.grid(cp = c(0.0010281))

# Expanded Regression Tree Model
epgc_mod_cart2 <-
  train(job_satisfaction ~ .,
        data=epcg_train3,  
        method = "rpart", 
        metric = "ROC", 
        tuneGrid = tune_cart2,
        trControl = control_conditions)

# Random Forest Model
epgc_mod_rf <-
  train(job_satisfaction ~ ., 
        data=epcg_train3,  
        method = "ranger", 
        metric = "ROC", 
        importance = 'impurity', 
        trControl = control_conditions)

# Set tuning parameters for Random Forest Model
epgc_tune_rf <- expand.grid(mtry=1:10, splitrule="gini", min.node.size=5)

# Tuned Random Forest Model
epgc_mod_rf2 <-
  train(job_satisfaction ~ ., 
        data=epcg_train3,  
        method = "ranger", 
        metric = "ROC", 
        importance = 'impurity',
        tuneGrid = epgc_tune_rf,
        trControl = control_conditions)

mod_list <-
  list(
    knn1 = epgc_mod_knn,
    knn2 = epgc_mod_knn2,
    cart1 = epgc_mod_cart,
    cart2 = epgc_mod_cart2,
    rf = epgc_mod_rf,
    rf2 = epgc_mod_rf2
  )
```

## Results - Machine Learning: Dot Plot Performance Measurement
```{r}
# Plotting Performance
dotplot(resamples(mod_list))
```

## Results - Machine Learning: Tuned Random Forest, Recent Graduates
```{r}
pred <- predict(epgc_mod_rf2, newdata = epcg_test3)
confusionMatrix(table(pred,epcg_test3$job_satisfaction))
```

## Results - Machine Learning: Tuned Random Forest Variable Importance
```{r, fig.height=10,fig.width=5, cache=TRUE}
plot(varImp(epgc_mod_rf2))
```

```{r, include=FALSE}
# Set up data and machine learning model for full sample
epcg17_all <-  epcg17 %>% 
  filter(., ugloanr!="01" & emrg<10) %>% 
  # selecting age, race, gender, undergraduate loan amoutn, graduate loan amount, physical disability indicator
  select(age, racethm, ugloanr, gender, grloanr, hcapin,
         #degree information, second to fifth highest degrees, highest degree, most recent degree
         d2dg, d3dg, d4dg, d5dg, dgrdg, mrdg,
         #primary/secondary work activity, public/private status of undergrad institution, year ba received
         actcap:acttch, bapbpr, bayr,
         # reason for changing employer
         chchg:chloc, chot:chsch,
         # visa type for non-US citizen, year highest degree awarded, father and mother's education level
         ctzfor, dgryr, eddad, edmom,
         #region code for employer, employer type, employer size, detailed employer sector, 
         emrg, emtp, emsize, emsecdt, emsecsm, facadv:facsoc,
         # hours per week worked, job benefits, job satisfaction, labor force status, looking for work, marital status
         hrswk, jobins:jobproft, jobvac, jobsatis, lfstat, lookwk, marsta,
         # job code for last job, job code for principal job, field of study of BA degree, employer type
         n2ocmlst, n2ocprmg, nbamed, nedtp,
         # reasons for working outside field of highest degree, most important reason for working outside field
         nrchg:nrpay, nrrea,
         # reasons for not working, extent job is related to highest degree, reason worked part-time
         nwfam:nwstu, ocedrlp, pjfam:pjstu,
         # respondent location, salary, satisfaction in prinicipal job for..., spouse working indicator, year principal job started
         resploc, salary, satadv:satsoc, spowk, strtyr,
         # amount still owed from undergrad loan, amount still owed from grad loan, veteran status 
         ugower, grower, vetstat) %>% 
  # creating gender and race indicators
  mutate(., female = ifelse(gender=="F", 1,0)) %>% 
  mutate(., race_white = ifelse(racethm=="5",1,0 )) %>% 
  mutate(., race_black = ifelse(racethm=="3",1,0 )) %>% 
  mutate(., race_hispanic = ifelse(racethm=="4",1,0 )) %>%
  mutate(., race_asian = ifelse(racethm=="1",1,0 )) %>%
  mutate(., race_other = ifelse(racethm=="2" | racethm=="6" | racethm=="7", 1,0)) %>% 
  # replacing 9999998 to NA for salary variable
  mutate(., salary= ifelse(salary==9999998, NA, salary)) %>% 
  # bachelor degree indicator
  mutate(., bachelor_degree= ifelse(mrdg=="1", 1,
                                    ifelse(dgrdg=="1", 1,
                                           ifelse(d2dg=="1", 1,
                                                  ifelse(d3dg=="1", 1,
                                                         ifelse(d4dg=="1",1,
                                                                ifelse(d5dg=="1", 1, 0))))))) %>%
  # master's degree indicator
  mutate(., master_degree= ifelse(mrdg=="2", 1,
                                    ifelse(dgrdg=="2", 1,
                                           ifelse(d2dg=="2", 1,
                                                  ifelse(d3dg=="2", 1,
                                                         ifelse(d4dg=="2",1,
                                                                ifelse(d5dg=="2", 1, 0))))))) %>%
  mutate(., doctorate_degree= ifelse(mrdg=="3", 1,
                                    ifelse(dgrdg=="3", 1,
                                           ifelse(d2dg=="3", 1,
                                                  ifelse(d3dg=="3", 1,
                                                         ifelse(d4dg=="3",1,
                                                                ifelse(d5dg=="3", 1, 0))))))) %>%
  # professional degree
  mutate(., professional_degree= ifelse(mrdg=="4", 1,
                                    ifelse(dgrdg=="4", 1,
                                           ifelse(d2dg=="4", 1,
                                                  ifelse(d3dg=="4", 1,
                                                         ifelse(d4dg=="4",1,
                                                                ifelse(d5dg=="4", 1, 0))))))) %>%
  mutate(.,satadv_satisfied = if_else(satadv=="1" | satadv =="2", 1, 0)) %>% 
  mutate(.,satben_satisfied = if_else(satben=="1" | satben =="2", 1, 0)) %>%
  mutate(.,satsal_satisfied = if_else(satsal=="1" | satsal =="2", 1, 0)) %>%
  mutate(.,satsec_satisfied = if_else(satsec=="1" | satsec =="2", 1, 0)) %>%
  mutate(.,satresp_satisfied = if_else(satresp=="1" | satresp =="2", 1, 0)) %>% 
  mutate(.,job_satisfied = if_else(jobsatis=="1" | jobsatis =="2", 1, 0))

# New machine learning data set with full observations
epcg_ml_all <- select(epcg17_all, "age"=age, "race"=racethm, "gender"=gender, "salary"=salary, 
                  "employer_size"=emsize, "employer_region"=emrg, "highest_degree"=dgrdg, 
                  "employer_sector"=emsecdt, "loan_amount"=ugloanr, "job_satisfaction"=job_satisfied)

# Ensure all factor levels are correct
epcg_ml_all$job_satisfaction <- factor(epcg_ml_all$job_satisfaction, 
                                      levels = c(1, 0), 
                                      labels = c("Satisfied", "Not_Satisfied"))

index = createDataPartition(epcg_ml_all$job_satisfaction,p=.75,list=F) 
epcg_train = epcg_ml_all[index,] # Use 75% of the data as training data 
epcg_test = epcg_ml_all[-index,] # holdout 25% as test data 

# Continuous variables that may need redistributing: salary
convert_salary <- . %>% mutate(salary = log(salary+1))

epcg_train2 <- epcg_train %>%  convert_salary()
epcg_test2 <- epcg_test %>%  convert_salary()

# Prep the recipe
rcp <- 
  recipe(job_satisfaction~.,epcg_train2) %>% 
  step_dummy(all_nominal(),-all_outcomes()) %>% 
  step_range(salary, age) %>%
  step_knnimpute(all_predictors()) %>%
  prep()

# Bake the recipe
epcg_train3 <- bake(rcp,epcg_train2)
epcg_test3 <- bake(rcp,epcg_test2)

# Set seed for replicability
set.seed(123)  

# Partition the data into 5 equal folds
folds <- createFolds(epcg_train3$job_satisfaction, k = 5) 

# Apply the folds
sapply(folds,length)

# Setting the control conditions to compare results across different model specifications
control_conditions <- 
  trainControl(method='cv',
               summaryFunction = twoClassSummary,
               classProbs = TRUE, 
               index = folds)
```

## Results - Machine Learning: Tuned Random Forest, Full Sample
```{r, cache=TRUE}
# Set tuning parameters for Random Forest Model
epgc_tune_rf <- expand.grid(mtry=1:10, splitrule="gini", min.node.size=5)

# Tuned Random Forest Model
epgc_mod_rf2 <-
  train(job_satisfaction ~ ., 
        data=epcg_train3,  
        method = "ranger", 
        metric = "ROC", 
        importance = 'impurity',
        tuneGrid = epgc_tune_rf,
        trControl = control_conditions)

pred_rf2_all <- predict(epgc_mod_rf2, newdata = epcg_test3)
confusionMatrix(table(pred_rf2_all,epcg_test3$job_satisfaction))
```

```{r, include=FALSE}
# Set up ML for Salary Satisfaction
epcg_ml_sal <- select(epcg17_new, "age"=age, "race"=racethm, "gender"=gender, "salary"=salary, 
                  "employer_size"=emsize, "employer_region"=emrg, "highest_degree"=dgrdg, 
                  "employer_sector"=emsecdt, "loan_amount"=ugloanr, "salary_satisfaction"=satsal_satisfied)

epcg_ml_sal$salary_satisfaction <- factor(epcg_ml_sal$salary_satisfaction, 
                                      levels = c(1, 0), 
                                      labels = c("Satisfied", "Not_Satisfied"))

index = createDataPartition(epcg_ml_sal$salary_satisfaction,p=.75,list=F) 
epcg_train = epcg_ml_sal[index,] # Use 75% of the data as training data 
epcg_test = epcg_ml_sal[-index,] # holdout 25% as test data 

# Continuous variables that may need redistributing: salary
convert_salary <- . %>% mutate(salary = log(salary+1))

epcg_train2 <- epcg_train %>%  convert_salary()
epcg_test2 <- epcg_test %>%  convert_salary()

# Prep the recipe
rcp <- 
  recipe(salary_satisfaction~.,epcg_train2) %>% 
  step_dummy(all_nominal(),-all_outcomes()) %>% 
  step_range(salary, age) %>%
  step_knnimpute(all_predictors()) %>%
  prep()

# Bake the recipe
epcg_train3 <- bake(rcp,epcg_train2)
epcg_test3 <- bake(rcp,epcg_test2)

# Set seed for replicability
set.seed(123)  

# Partition the data into 5 equal folds
folds <- createFolds(epcg_train3$salary_satisfaction, k = 5) 

# Apply the folds
sapply(folds,length)

# Setting the control conditions to compare results across different model specifications
control_conditions <- 
  trainControl(method='cv',
               summaryFunction = twoClassSummary,
               classProbs = TRUE, 
               index = folds)

# Set tuning parameters for Random Forest Model
epgc_tune_rf <- expand.grid(mtry=1:10, splitrule="gini", min.node.size=5)
```

## Results - Machine Learning: Tuned Random Forest, Salary Satisfaction
```{r, cache=TRUE}
# Tuned Random Forest Model
epgc_mod_rf2 <-
  train(salary_satisfaction ~ ., 
        data=epcg_train3,  
        method = "ranger", 
        metric = "ROC", 
        importance = 'impurity',
        tuneGrid = epgc_tune_rf,
        trControl = control_conditions)

pred_rf2_salary <- predict(epgc_mod_rf2, newdata = epcg_test3)
confusionMatrix(table(pred_rf2_salary,epcg_test3$salary_satisfaction))
```

```{r}
# Set up ML for Career Advancement Satisfaction
epcg_ml_adv <- select(epcg17_new, "age"=age, "race"=racethm, "gender"=gender, "salary"=salary,
                  "employer_size"=emsize, "employer_region"=emrg, "highest_degree"=dgrdg, 
                  "employer_sector"=emsecdt, "loan_amount"=ugloanr, "adv_satisfaction"=satadv_satisfied)

epcg_ml_adv$adv_satisfaction <- factor(epcg_ml_adv$adv_satisfaction, 
                                      levels = c(1, 0), 
                                      labels = c("Satisfied", "Not_Satisfied"))

index = createDataPartition(epcg_ml_adv$adv_satisfaction,p=.75,list=F) 
epcg_train = epcg_ml_adv[index,] # Use 75% of the data as training data 
epcg_test = epcg_ml_adv[-index,] # holdout 25% as test data 

# Continuous variables that may need redistributing: salary
convert_salary <- . %>% mutate(salary = log(salary+1))

epcg_train2 <- epcg_train %>%  convert_salary()
epcg_test2 <- epcg_test %>%  convert_salary()

# Prep the recipe
rcp <- 
  recipe(adv_satisfaction~.,epcg_train2) %>% 
  step_dummy(all_nominal(),-all_outcomes()) %>% 
  step_range(salary, age) %>%
  step_knnimpute(all_predictors()) %>%
  prep()

# Bake the recipe
epcg_train3 <- bake(rcp,epcg_train2)
epcg_test3 <- bake(rcp,epcg_test2)

# Set seed for replicability
set.seed(123)  

# Partition the data into 5 equal folds
folds <- createFolds(epcg_train3$adv_satisfaction, k = 5) 

# Apply the folds
sapply(folds,length)

# Setting the control conditions to compare results across different model specifications
control_conditions <- 
  trainControl(method='cv',
               summaryFunction = twoClassSummary,
               classProbs = TRUE, 
               index = folds)

# Set tuning parameters for Random Forest Model
epgc_tune_rf <- expand.grid(mtry=1:10, splitrule="gini", min.node.size=5)
```

## Results - Machine Learning: Tuned Random Forest, Career Advancement Satisfaction
```{r, cache=TRUE}
# Tuned Random Forest Model
epgc_mod_rf2 <-
  train(adv_satisfaction ~ ., 
        data=epcg_train3,  
        method = "ranger", 
        metric = "ROC", 
        importance = 'impurity',
        tuneGrid = epgc_tune_rf,
        trControl = control_conditions)

pred_rf2_adv <- predict(epgc_mod_rf2, newdata = epcg_test3)
confusionMatrix(table(pred_rf2_adv,epcg_test3$adv_satisfaction))
```

## Limitations

1. Low variation in outcome variable resulted in low sensitivity
2. Satisfaction may not perfectly represent job mismatch
3. Loan amount categories instead of numerical values

## Conclusions

1. Difficult to predict job mismatch and satisfaction given current data
2. Salary is the most important predictor job satisfaction
3. Simple to predict job satisfaction, though potential response biases limit our ability to draw conclusions from this finding




