---
title: "Predicting Job Mismatch from Job Satisfaction"
author: "Harri-Lee-Sun"
date: "2019/4/27"
output:
  ioslides_presentation:
    incremental: yes
    widescreen: yes
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Problem statement and background

- The cost of higher education has increased but incomes hasn't

- Student debt has also increased significantly

- Student loan debt in 2018: $1.5 trillion

- Graduating class of 2016 has average of $37,000 student loan.

## Problem statement and background

- Studies have shown that student debt impacts life decisions

- Student debt may pressure and lead to job mismatches

- Hypothesis: high levels of student loan debt have negative impacts on employment outcomes

## Research design and approach

- Goals

1. Data visualization: to test any trends or relationships between variables of interest

2. Machine learning: to find the best predictor for the overall job satisfaction

## Methods / Tools

- Dataset: National Survey of College Graduates (NSCG)

- Data Visualization 
1. Job satisfaction over Loan Amounts (Region and Race)
2. Comparison of job satisfaction aspects
3. Distribution of job satisfaction responses by salary and age groups

## Methods / Tools

- Machine Learning 
1.
2.
3.


## Results - Visualization 1

```{r, include=FALSE}
# Load Packages
library(haven)
library(tidyverse)
library(caret)
library(recipes)
library(pdp)
library(skimr)
library(foreach)
library(pastecs)
library(ggplot2)

## Data Wrangling - Won't be included in the knit pdf output.
## filtering by age, selecting variables of interest, creating key indicators
epcg17 <- read_dta("epcg17.dta")
epcg17_new <-  epcg17 %>% 
  filter(., age<36 & bast=="099" & bayr>1975 & ugloanr!="01" & emrg<10) %>% 
  # selecting age, race, gender, undergraduate loan amoutn, graduate loan amount, physical disability indicator
  select(age, racethm, ugloanr, gender, grloanr, hcapin,
         #degree information, second to fifth highest degrees, highest degree, most recent degree
         d2dg, d3dg, d4dg, d5dg, dgrdg, mrdg,
         #primary/secondary work activity, public/private status of undergrad institution, year ba received
         actcap:acttch, bapbpr, bayr,
         # reason for changing employer
         chchg:chloc, chot:chsch,
         # visa type for non-US citizen, year highest degree awarded, father and mother's education level
         ctzfor, dgryr, eddad, edmom,
         #region code for employer, employer type, employer size, detailed employer sector, 
         emrg, emtp, emsize, emsecdt, emsecsm, facadv:facsoc,
         # hours per week worked, job benefits, job satisfaction, labor force status, looking for work, marital status
         hrswk, jobins:jobproft, jobvac, jobsatis, lfstat, lookwk, marsta,
         # job code for last job, job code for principal job, field of study of BA degree, employer type
         n2ocmlst, n2ocprmg, nbamed, nedtp,
         # reasons for working outside field of highest degree, most important reason for working outside field
         nrchg:nrpay, nrrea,
         # reasons for not working, extent job is related to highest degree, reason worked part-time
         nwfam:nwstu, ocedrlp, pjfam:pjstu,
         # respondent location, salary, satisfaction in prinicipal job for..., spouse working indicator, year principal job started
         resploc, salary, satadv:satsoc, spowk, strtyr,
         # amount still owed from undergrad loan, amount still owed from grad loan, veteran status 
         ugower, grower, vetstat) %>% 
  # creating gender and race indicators
  mutate(., female = ifelse(gender=="F", 1,0)) %>% 
  mutate(., race_white = ifelse(racethm=="5",1,0 )) %>% 
  mutate(., race_black = ifelse(racethm=="3",1,0 )) %>% 
  mutate(., race_hispanic = ifelse(racethm=="4",1,0 )) %>%
  mutate(., race_asian = ifelse(racethm=="1",1,0 )) %>%
  mutate(., race_other = ifelse(racethm=="2" | racethm=="6" | racethm=="7", 1,0)) %>% 
  # replacing 9999998 to NA for salary variable
  mutate(., salary= ifelse(salary==9999998, NA, salary)) %>% 
  # bachelor degree indicator
  mutate(., bachelor_degree= ifelse(mrdg=="1", 1,
                                    ifelse(dgrdg=="1", 1,
                                           ifelse(d2dg=="1", 1,
                                                  ifelse(d3dg=="1", 1,
                                                         ifelse(d4dg=="1",1,
                                                                ifelse(d5dg=="1", 1, 0))))))) %>%
  # master's degree indicator
  mutate(., master_degree= ifelse(mrdg=="2", 1,
                                    ifelse(dgrdg=="2", 1,
                                           ifelse(d2dg=="2", 1,
                                                  ifelse(d3dg=="2", 1,
                                                         ifelse(d4dg=="2",1,
                                                                ifelse(d5dg=="2", 1, 0))))))) %>%
  mutate(., doctorate_degree= ifelse(mrdg=="3", 1,
                                    ifelse(dgrdg=="3", 1,
                                           ifelse(d2dg=="3", 1,
                                                  ifelse(d3dg=="3", 1,
                                                         ifelse(d4dg=="3",1,
                                                                ifelse(d5dg=="3", 1, 0))))))) %>%
  # professional degree
  mutate(., professional_degree= ifelse(mrdg=="4", 1,
                                    ifelse(dgrdg=="4", 1,
                                           ifelse(d2dg=="4", 1,
                                                  ifelse(d3dg=="4", 1,
                                                         ifelse(d4dg=="4",1,
                                                                ifelse(d5dg=="4", 1, 0))))))) %>%
  mutate(.,satadv_satisfied = if_else(satadv=="1" | satadv =="2", 1, 0)) %>% 
  mutate(.,satben_satisfied = if_else(satben=="1" | satben =="2", 1, 0)) %>%
  mutate(.,satsal_satisfied = if_else(satsal=="1" | satsal =="2", 1, 0)) %>%
  mutate(.,satsec_satisfied = if_else(satsec=="1" | satsec =="2", 1, 0)) %>%
  mutate(.,satresp_satisfied = if_else(satresp=="1" | satresp =="2", 1, 0)) %>% 
  mutate(.,job_satisfied = if_else(jobsatis=="1" | jobsatis =="2", 1, 0))
```
```{r}
## Visual 2
# data set creation for the visual 2
dataforplot2 <- epcg17_new %>%
  filter(., jobsatis!="L") %>%
  select(., ugloanr,satadv_satisfied:satresp_satisfied,job_satisfied,emrg) %>% 
  mutate(.,obscount=1) %>%
  mutate(.,ugloanr_group=ifelse(as.numeric(ugloanr)==2, 1,
                                ifelse(as.numeric(ugloanr)>2 & as.numeric(ugloanr)<6,2,
                                       ifelse(as.numeric(ugloanr)>5 & as.numeric(ugloanr)<9, 3,
                                              ifelse(as.numeric(ugloanr)>8 &  as.numeric(ugloanr)<12,4,5))))) %>%
  group_by(.,ugloanr_group, emrg) %>% 
  summarize(.,mean_job_sat=mean(job_satisfied),
            obscount=sum(obscount)) %>% 
  mutate(.,emrg_char=ifelse(emrg=="01","New England",
                            ifelse(emrg=="02", "Middle Atlantic",
                                   ifelse(emrg=="03", "East North Central",
                                          ifelse(emrg=="04", "West North Central",
                                                 ifelse(emrg=="05", "South Central",
                                                        ifelse(emrg=="06", "East South Central",
                                                               ifelse(emrg=="07", "West South Central",
                                                                      ifelse(emrg=="08", "Mountain", "Pacific and US Territories"))))))))) %>% 
  filter(., obscount>=50) %>% 
  mutate(.,emrg_factor=factor(emrg_char, c("New England", "Middle Atlantic", "East North Central", "West North Central",
                                            "South Central", "East South Central", "West South Central",
                                            "Mountain", "Pacific and US Territories"))) %>%
  mutate(.,mean_job_sat_100 = mean_job_sat * 100)

# ggplot code for visual 2
ggplot(dataforplot2, aes(x=emrg_factor, y=mean_job_sat_100, color=as.character(ugloanr_group), group=as.character(ugloanr_group))) +
  geom_point(size=3.5, alpha=.8) +
  geom_text(aes(label=round(mean_job_sat_100,0)), size=4, vjust = -.8, nudge_y = 0.05) +
  scale_color_manual(name = "Undergrad Loan Amount",
                     values = c("1"="yellowgreen", "2"="yellow3", "3"="orange1", "4"="orangered1", "5" = "red4"),
                     labels = c("1"="$0", "2"="$1-30,000", "3"="$30,001-60,000", "4"="$60,001-90,000", "5" = "More than $90,000")) +
  labs(title ="Percent Satisfied with Job by Undergraduate Loan Amount and Region",
       caption = "Data: National Survey of College Graduates") +
  coord_flip() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(),
        panel.grid.major.y =element_line(colour = "grey", size = 0.5))

```


## Results - Visualization 2
```{r}
## Results - Visualization
# Visual 4 data preparation
dataforplot4 <- epcg17_new %>%
  filter(., jobsatis!="L") %>%
  select(., ugloanr,emrg, satadv_satisfied:satresp_satisfied, job_satisfied) %>% 
  mutate(.,obscount=1) %>%
  mutate(.,ugloanr_group=ifelse(as.numeric(ugloanr)==2, 1,
                                ifelse(as.numeric(ugloanr)>2 & as.numeric(ugloanr)<6,2,
                                       ifelse(as.numeric(ugloanr)>5 & as.numeric(ugloanr)<9, 3,
                                              ifelse(as.numeric(ugloanr)>8 & as.numeric(ugloanr)<12,4,5))))) %>%
  group_by(.,ugloanr_group, emrg) %>%
  summarize(., mean1=mean(satadv_satisfied),
            mean2=mean(satben_satisfied), mean3=mean(satsal_satisfied),
            mean4=mean(satsec_satisfied), mean5=mean(satresp_satisfied),
            mean6=mean(job_satisfied), obscount=sum(obscount)) %>%
  filter(., obscount>=50) %>% 
  ungroup(.,ugloanr_group, emrg) %>% 
  gather(., key, value,mean1:mean6) %>% 
  mutate(.,emrg_char=ifelse(emrg=="01","New England",
                            ifelse(emrg=="02", "Middle Atlantic",
                                   ifelse(emrg=="03", "East North Central",
                                          ifelse(emrg=="04", "West North Central",
                                                 ifelse(emrg=="05", "South Central",
                                                        ifelse(emrg=="06", "East South Central",
                                                               ifelse(emrg=="07", "West South Central",
                                                                      ifelse(emrg=="08", "Mountain", "Pacific and US Territories"))))))))) %>% 
  mutate(.,emrg_factor=factor(emrg_char, c("New England", "Middle Atlantic", "East North Central", "West North Central",
                                           "South Central", "East South Central", "West South Central",
                                           "Mountain", "Pacific and US Territories"))) %>% 
  mutate(.,satisfaction_type=ifelse(key=="mean1","Career Advancement",
                                    ifelse(key=="mean2", "Benefits",
                                           ifelse(key=="mean3", "Salary",
                                                  ifelse(key=="mean4", "Job Security",
                                                         ifelse(key=="mean5", "Responsibility", "Job - Overall")))))) %>% 
  mutate(.,mean100=value*100)

## Visual 4
ggplot(dataforplot4, aes(x=as.character(ugloanr_group),y=mean100, color=satisfaction_type, group=satisfaction_type)) +
  geom_point() +
  geom_line(size=1.3, alpha=.6) +
  facet_wrap(~emrg_factor) +
  scale_color_brewer(name="Satisfaction Category",
                     type="qual",
                     palette="Set1") +
  scale_x_discrete(labels = c("1"="$0", "2"="$1-30K", "3"="$30k-60k", "4"="$60k-90k", "5" = "More than $90k")) +
  scale_y_continuous(limits = c(60, 100),breaks = seq(60, 100, 10)) +
  labs(title ="Types of Job Satisfaction by Undergraduate Loan Amounts and by Region",
       caption = "Data: National Survey of College Graduates",
       y = "Percent Satisfied") +
  theme(axis.title.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(),
        panel.grid.major.y =element_line(colour = "grey", size = 0.5),
        axis.text.x = element_text(size=8, angle=20))
```

## Results - Machine Learning
```{r}

```
## Limitations

1. Low variation in outcome variable resulted in low sensitivity
2. Satisfaction may not perfectly represent job mismatch
3. Loan amount categories instead of numerical values

## Next Steps

1. Designing the survey to measure job mismatch
2. 




