---
title: "Predicting job mismatch through analyzing overall job satisfaction"
author: "Harri-Lee-Sun"
date: "2019/4/27"
output:
  ioslides_presentation:
    incremental: yes
    widescreen: yes
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Problem statement and background

- The cost of higher education has increased but incomes hasn't

- Student debt has also increased significantly

- Student loan debt in 2018: $1.5 trillion

- Graduating class of 2016 has average of $37,000 student loan.

## Problem statement and background

- Less studies focus on the effects of the level of students loans debt has on students’ post-graduation decision-making

- We hypothesize that high levels of student loan debt have negative effects on employment outcomes for recent graduates

## Research design and approach

- Dataset: National Survey of College Graduates (NSCG)

- Goals
1. Data visualization: to test any trends or relationships between variables of interest
2. Machine learning: to find the best predictor for the overall job satisfaction

## Methods / tools

- Data Visualization 
1. recode variables
2. create new indicators
3. group variables 
- Machine Learning 



## Slide with Plot

```{r pressure}
plot(pressure)
```



## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```


