---
title: "Predicting job mismatch through analyzing overall job satisfaction"
author: "Harri-Lee-Sun"
date: "2019/4/27"
output:
  ioslides_presentation:
    incremental: yes
    widescreen: yes
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Problem statement and background

- The cost of higher education has increased but incomes hasn't

- Student debt has also increased significantly

- Student loan debt in 2018: $1.5 trillion

- Graduating class of 2016 has average of $37,000 student loan.

## Problem statement and background

- Studies have shown that student debt impacts life decisions

- Student debt may pressure and lead to job mismatches

- Hypothesis: high levels of student loan debt have negative impacts on employment outcomes

## Research design and approach

- Goals

1. Data visualization: to test any trends or relationships between variables of interest

2. Machine learning: to find the best predictor for the overall job satisfaction

## Methods / Tools

- Dataset: National Survey of College Graduates (NSCG)

- Data Visualization 
1. Job satisfaction over Loan Amounts (Region and Race)
2. Comparison of job satisfaction aspects
3. Distribution of job satisfaction responses by salary and age groups

## Methods / Tools

- Machine Learning 
1.
2.
3.


## Results - Visualization

```{r pressure}
plot(pressure)
```



## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```


